On exponential sums arising from the classical doubling method

### 2025-03-01 paper.tex
Epsilon no more

\begin{theorem}\label{thm:computation-of-doubling-gauss-sum-scalar-for-deligne-lusztig-characters}
	Suppose that $T \cong \prod_{j=1}^r \multiplicativegroup{\quadraticFieldExtension{k_j}} \times \prod_{i=1}^s \finiteFieldExtension{2m_i}^1$ is a maximal torus of $G$. Suppose that $\theta \colon T \to \multiplicativegroup{\cComplex}$ is a character and that under the above isomorphism $\theta = \alpha_1 \times \dots \times \alpha_r \times \theta_1 \times \dots \times \theta_s \times \epsilon$, where $\alpha_j \colon \multiplicativegroup{\quadraticFieldExtension{k_j}} \to \multiplicativegroup{\cComplex}$ and $\theta_i \colon \finiteFieldExtension{2m_i}^1 \to \multiplicativegroup{\cComplex}$ are characters for every $i$ and $j$. Denote $$g_T\left(\chi, \theta, \fieldCharacter\right) = \prod_{i=1}^s \GaussSumCharacter{\theta'_i}{\chi^{-1}}{\fieldCharacter_{2m_i}} \cdot  \prod_{j=1}^r \GaussSumCharacter{\alpha_j}{\chi^{-1}}{\fieldCharacter_{k_j}} \GaussSumCharacter{\alpha_j^{-1}}{\chi^{-1}}{\fieldCharacter_{k_j}}.$$ Then
	\begin{align*}
		 \dblVirtualGaussSumScalar{R_{T, \theta}}{\chi} =& R_{T,\theta}\left(1\right) \GaussSumScalar{\chi^2}{\fieldCharacter}^{\left\lfloor\frac{\dim_{\quadraticExtension} \hermitianSpace}{2}\right\rfloor} g_T\left(\chi, \theta, \fieldCharacter\right) \prod_{i=1}^s \theta_i\left(-1\right) \cdot \prod_{j=1}^r \alpha_j\left(-1\right).
	\end{align*}
\end{theorem}
\begin{proof}
	This follows from the results mentioned above and from the formula $$R_T\left(1\right) = \varepsilon_{\algebraicGroup{G}} \varepsilon_{\algebraicGroup{T}} \frac{\grpIndex{G}{T}}{\sizeof{G}_p},$$ the fact that $$ \sizeof{\lieAlgebra} = \sizeof{G}_p^2 \cdot q^{\left\lfloor\frac{\dim_{\quadraticExtension} \hermitianSpace}{2}\right\rfloor},$$
	the facts \begin{equation*}
		\varepsilon_{\algebraicGroup{T}} = \left(-1\right)^r \text{\,\,\,\,\,\, and \,\,\,\,\,\,}  \varepsilon_{\algebraicGroup{G}} = \left(-1\right)^{\left\lfloor\frac{\dim_{\quadraticExtension} \hermitianSpace}{2}\right\rfloor},
	\end{equation*}
	and finally from $$\GaussSumScalar{\chi^2}{\fieldCharacter} = -\tau\left(\chi^2, \fieldCharacter\right).$$
\end{proof}
\begin{remark}
	Suppose that $\pi$ is any irreducible representation of $G$ such that $\trace \pi$ appears with non-zero coefficient in $R_{T,\theta}$. Then the central character of $\pi$ satisfies
	\begin{align*}
		\centralCharacter{\pi}\left(-1\right) = \theta\left(-1\right) =& \prod_{j=1}^r \alpha_j\left(-1\right) \cdot \prod_{i=1}^s \theta_i\left(-1\right).
	\end{align*}
	Thus \begin{align*}
		\dblVirtualGaussSumScalar{R_{T, \theta}}{\chi} =& R_{T,\theta}\left(1\right) \GaussSumScalar{\chi^2}{\fieldCharacter}^{\left\lfloor\frac{\dim_{\quadraticExtension} \hermitianSpace}{2}\right\rfloor} \centralCharacter{\pi}\left(-1\right) g_T\left(\theta, \fieldCharacter\right).
	\end{align*}
\end{remark}

>>>>

\begin{theorem}\label{thm:computation-of-doubling-gauss-sum-scalar-for-deligne-lusztig-characters}
	Suppose that $T \cong \prod_{j=1}^r \multiplicativegroup{\quadraticFieldExtension{k_j}} \times \prod_{i=1}^s \finiteFieldExtension{2m_i}^1$ is a maximal torus of $G$. Suppose that $\theta \colon T \to \multiplicativegroup{\cComplex}$ is a character and that under the above isomorphism $\theta = \alpha_1 \times \dots \times \alpha_r \times \theta_1 \times \dots \times \theta_s$, where $\alpha_j \colon \multiplicativegroup{\quadraticFieldExtension{k_j}} \to \multiplicativegroup{\cComplex}$ and $\theta_i \colon \finiteFieldExtension{2m_i}^1 \to \multiplicativegroup{\cComplex}$ are characters for every $i$ and $j$. Denote $$g_T\left(\chi, \theta, \fieldCharacter\right) = \prod_{i=1}^s \GaussSumCharacter{\theta'_i}{\chi^{-1}}{\fieldCharacter_{2m_i}} \cdot  \prod_{j=1}^r \GaussSumCharacter{\alpha_j}{\chi^{-1}}{\fieldCharacter_{k_j}} \GaussSumCharacter{\alpha_j^{-1}}{\chi^{-1}}{\fieldCharacter_{k_j}}.$$ Then
	\begin{align*}
		 \dblVirtualGaussSumScalar{R_{T, \theta}}{\chi} =& R_{T,\theta}\left(1\right) \GaussSumScalar{\chi^2}{\fieldCharacter}^{\left\lfloor\frac{\dim_{\quadraticExtension} \hermitianSpace}{2}\right\rfloor} g_T\left(\chi, \theta, \fieldCharacter\right) \prod_{i=1}^s \theta_i\left(-1\right) \cdot \prod_{j=1}^r \alpha_j\left(-1\right).
	\end{align*}
\end{theorem}
\begin{proof}
	This follows from the results mentioned above and from the formula $$R_T\left(1\right) = \varepsilon_{\algebraicGroup{G}} \varepsilon_{\algebraicGroup{T}} \frac{\grpIndex{G}{T}}{\sizeof{G}_p},$$ the fact that $$ \sizeof{\lieAlgebra} = \sizeof{G}_p^2 \cdot q^{\left\lfloor\frac{\dim_{\quadraticExtension} \hermitianSpace}{2}\right\rfloor},$$
	the facts \begin{equation*}
		\varepsilon_{\algebraicGroup{T}} = \left(-1\right)^r \text{\,\,\,\,\,\, and \,\,\,\,\,\,}  \varepsilon_{\algebraicGroup{G}} = \left(-1\right)^{\left\lfloor\frac{\dim_{\quadraticExtension} \hermitianSpace}{2}\right\rfloor},
	\end{equation*}
	and finally from $$\GaussSumScalar{\chi^2}{\fieldCharacter} = -\tau\left(\chi^2, \fieldCharacter\right).$$
\end{proof}
\begin{remark}
	Suppose that $\pi$ is any irreducible representation of $G$ such that $\trace \pi$ appears with non-zero coefficient in $R_{T,\theta}$. Then the central character of $\pi$ satisfies
	\begin{align*}
		\centralCharacter{\pi}\left(-1\right) = \theta\left(-1\right) =& \prod_{j=1}^r \alpha_j\left(-1\right) \cdot \prod_{i=1}^s \theta_i\left(-1\right).
	\end{align*}
	Thus \begin{align*}
		\dblVirtualGaussSumScalar{R_{T, \theta}}{\chi} =& R_{T,\theta}\left(1\right) \GaussSumScalar{\chi^2}{\fieldCharacter}^{\left\lfloor\frac{\dim_{\quadraticExtension} \hermitianSpace}{2}\right\rfloor} \centralCharacter{\pi}\left(-1\right) g_T\left(\theta, \fieldCharacter\right).
	\end{align*}
\end{remark}
